# Interviewing data: R, RStudio, and the tidyverse for data analysis

```{r setup, include=FALSE}
knitr::opts_chunk$set(out.width="900px", out.height="500px", dpi=300, message=FALSE, warning=FALSE)
knitr::opts_knit$set(root.dir = "../data/r-analysis")
```

### R and RStudio

Last quarter we worked with **[R](http://www.r-project.org/)** and **[RStudio](https://www.rstudio.com/)** to process data and make static, interactive, and animated graphics. Here, we will explore how to "interview" data to find leads and context for stories.

### The data we will use

Download the data for these classes from [here](data/r-analysis.zip), unzip the folder and place it on your desktop. It contains the following files:

- `ca_discipline.csv` Disciplinary alerts and actions issued by the Medical Board of California from 2008 to 2017. Processed from downloads available [here](https://www.mbc.ca.gov/Publications/Disciplinary_Actions/). Contains the following variables:

	- `alert_date` Date alert issued.
	- `last_name` Last name of doctor/health care provider.
	- `first_name` First name of doctor/health care provider.
	- `middle_name` Middle/other names.
	- `name_suffix` Name suffix (Jr., II etc)
	- `city` City of practive location.
	- `state` State of practice location.
	- `license` California medical license number.
	- `action_type` Type of action.
	- `action_date` Date of action.


- `ca_medicare_opioids.csv` Data on prescriptions of opioid drugs under the Medicare Part D Prescription Drug Program by doctors in California, from 2013 to 2015. Filtered from the national data downloads available [here](https://www.cms.gov/Research-Statistics-Data-and-Systems/Statistics-Trends-and-Reports/Medicare-Provider-Charge-Data/Part-D-Prescriber.html). It contains many variables, including:
  - `npi` [National Provider Identifier](https://npiregistry.cms.hhs.gov/) (NPI) for the doctor/organization making the claim. This is a unique code for each health care provider.
  - `nppes_provider_last_org_name` For individual doctors, their last name. For organizations, the organziation name.
  - `nppes_provider_first_name` First name for individual doctors, blank for organizations.
  - `nppes_provider_city` City where the provider is located.
  - `nppes_provider_state` State where the provider is located; "CA" for all of these records.
  - `specialty_description` Provider's medical speciality, reported on their medicare claims. For providers that have more than one Medicare specialty code reported on their claims, the code associated with the largest number of services.
  - `drug_name` Includes both brand names (drugs that have a trademarked name) and generic names (drugs that do not have a trademarked name).
  - `generic_name` The chemical ingredient of a drug rather than the trademarked brand name under which the drug is sold.
  - `bene_count` Total number of unique Medicare Part D beneficiaries (i.e. patients) with at least one claim for the drug. Counts fewer than 11 are suppressed and are indicated by a blank.
  - `total_claim_count` Number of Medicare Part D claims; includes original prescriptions and refills. If less than 11, counts are not included in the data file.
  - `total_drug_cost` Total cost paid for all associated claims; includes ingredient cost, dispensing fee, sales tax, and any applicable fees.
  - `year` 2013, 2014, or 2015.

- `npi_license.csv` Crosswalk file to join NPI identifiers to state license numbers, processed from the download available [here](https://www.nber.org/data/npi-state-license-crosswalk.html) to include license numbers potentially matching California doctors. This will provide one way of joining the precription data to the medical board disciplinary data. As we shall see, problems with the data mean that it is not infallible. Contains the following variables:
 - `npi` National Provider Identifier, as described above.
 - `plicnum` State license number, from the original file.
 - `license` Processed from `pclicnum` to conform to the format of California medical license numbers.

### Some words of caution, before we start

The US is currently in the grip of an epidemic of opioid abuse and addiction. Although [widespread medical prescription](https://www.buzzfeed.com/danvergano/2-old-painkiller-papers) of opioids [helped drive addiction](https://www.buzzfeed.com/danvergano/whats-causing-the-opioid-crisis), a [majority of overdoses](https://www.buzzfeed.com/danvergano/life-expectancy-opioid-overdoses) now occur through the consumption of drugs purchased illegally.

Opioids have important medical uses, and just because a doctor prescribes large amounts of the drugs doesn't necessarily mean they are practising irresponsibly.

As ProPublica explained, in [the methods](https://www.propublica.org/article/how-we-analyzed-medicares-drug-data-long-methodology) for its stories based on Medicare Part D prescription data:

> The data could not tell us everything. We interviewed many high-volume prescribers to better understand their patients and their practices. Some told us their numbers were high because they were credited with prescriptions by others working in the same practice. In addition, providers who primarily work in long-term care facilities or busy clinics with many patients naturally may write more prescriptions.

### Reminder on R code basics

We reviewed these basics in the previous quarter:

- `<-` is known as an “assignment operator.” It means: “Make the object named to the left equal to the output of the code to the right.”
- `&` means AND, in Boolean logic.
- `|` means OR, in Boolean logic.
- `!` means NOT, in Boolean logic.
- When referring to values entered as text, or to dates, put them in quote marks, like this: `"United States"`, or `"2016-07-26"`. Numbers are not quoted.
- When entering two or more values as a list, combine them using the function `c`, for combine, with the values separated by commas, for example: `c("2017-07-26","2017-08-04")`
- As in a spreadsheet, you can specify a range of values with a colon, for example: `c(1:10)` creates a list of integers (whole numbers) from one to ten.
- Some common operators:
 - `+` `-` add, subtract.
 -  `*` `/` multiply, divide.
 -  `>` `<` greater than, less than.
 -  `>=` `<=` greater than or equal to, less than or equal to.
 -  `!=` not equal to.

- Equals signs can be a little confusing, but see how they are used in the code we use today:

 -  `==` test whether an object is equal to a value. This is often used when filtering data, as we will see.
 -  `=` make an object equal to a value; similar to `<-`, but used within a function (see below).

- Handling null values:
 - Nulls are designated as `NA`.
 - `is.na(x)` looks for nulls within variable `x`.
 - `!is.na(x)` looks for non-null values within variable `x`.

Here, `is.na` is a **function**. Functions are followed by parentheses, and act on the data/code in the parentheses.

**Important:** Object and variable names in R should not contain spaces.

### Reminder on dplyr basics 

We can use the **[dplyr](https://dplyr.tidyverse.org/)** package process and analyze data, using these basic operations:

- **Sort:** Largest to smallest, oldest to newest, alphabetical etc.

- **Filter:** Select a defined subset of the data.

- **Summarize/Aggregate:** Deriving one value from a series of other values to produce a summary statistic. Examples include: count, sum, mean, median, maximum, minimum etc. Often you'll **group** data into categories first, and then aggregate by group.

- **Join:** Merging entries from two or more datasets based on common field(s), e.g. unique ID number, last name and first name.

Here, again, are some of the most useful functions in **dplyr**:

- `select` Choose which columns to include.
- `filter` **Filter** the data.
- `arrange` **Sort** the data, by size for continuous variables, by date, or alphabetically.
- `group_by` **Group** the data by a categorical variable.
- `summarize` **Summarize**, or aggregate (for each group if following `group_by`). Often used in conjunction with functions including:
	- `mean(x)` Calculate the mean, or average, for variable `x`.
	- `median(x)` Calculate the median.
	- `max(x)` Find the maximum value.
	- `min(x)` Find the minimum value.
	- `sum(x)` Add all the values together.
	- `n()` Count the number of records. Here there isn't a variable in the brackets of the function, because the number of records applies to all variables.
	- `n_distinct(x`) Count the number of unique values in variable `x`.
- `mutate` Create new column(s) in the data, or change existing column(s).
- `rename` Rename column(s).
- `bind_rows` Merge two data frames into one, combining data from columns with the same name.

There are a number of **join** functions in **dplyr** to combine data from two data frames. Here are the most useful:

- `inner_join` returns values from both tables only where there is a match.
- `left_join` returns all the values from the first-mentioned table, plus those from the second table that match.
- `semi_join` filters the first-mentioned table to include only values that have matches in the second table.
- `anti_join` filters the first-mentioned table to include only values that have no matches in the second table.

[Here is a useful reference](https://stat545-ubc.github.io/bit001_dplyr-cheatsheet.html) for managing joins with **dplyr**.

### Setting up

OpenRStudio, save a new script in the same folder as the downloaded data and set the working directory to this folder by selecting from the top menu `Session>Set Working Directory>To Source File Location`. (Doing so means we can load the files in this directory without having to refer to the full path for their location, and anything we save will be written to this folder.) 

Also install the [DT](https://rstudio.github.io/DT/) package, which turns R data frames into searchable web tables.

Now load the packages we'll use today:

```{r}
# load packages to read and write csv files, process data, work with dates, and make searchable web tables
library(readr)
library(dplyr)
library(lubridate)
library(DT)
```

At this point, and at regular intervals, save your script, by clicking the save/disk icon in the script panel, or using the `⌘-S`/`Ctrl-S` keyboard shortcut.

### Load and examine the disciplinary actions data


```{r}
# load ca medical board disciplinary actions data
ca_discipline <- read_csv("ca_discipline.csv")

# view structure of data
glimpse(ca_discipline)
```

#### Filter, sort, and clean data

To get used to working with **dplyr**, we will now start filtering and sorting the data according to the doctors' locations, and the types of disciplinary actions they faced.

First, let's look at the types of disciplinary actions in the data:

```{r}
# look at types of disciplinary actions
types <- ca_discipline %>%
  select(action_type) %>%
  unique()
```

This code first copies the `ca_disciplinary` data frame into a new object called `types`. Then (`%>%`) it `select`s the `action_type` variable only. Finally, it uses a function called `unique` to display the unique values in that variable, with no duplicates. This should be the result:

Now make a searchable web table to display the data:

```{r}
# make a searchable web table
datatable(types)
```

Many of the values for `action_types` seem to be subtle variations of the same thing. If we were going to all analyze the types in detail, we may need to group some of these together, after speaking to an expert to understand what they all mean. If you search the table for "reprimand"", however, it's pretty obvious that the four entries containing that word all amount to the same thing. To get a sense of how we could write code to clean up these categories, let's convert those to all say "Reprimand" :

```{r}
# clean data for reprimands
ca_discipline <- ca_discipline %>%
  mutate(action_type = case_when(grepl("reprimand", ignore.case = TRUE, action_type) ~ "Reprimand",
                                 TRUE ~ action_type))
```

The `case_when` function allows you to set up conditional statements. Here we are changing the `action_type` variable so that if the text contains the word "reprimand," matched using the `grepl` function, irrespective of the case of the letters, the entry is changed to "Reprimand," otherwise (the `TRUE` condition) it is left with the original `action_type` value. You can strong together multiple conditional statements in a `case_when` function to edit the data as you wish. 

Some of the action types are clear and unambiguous: `Revoked` is the medical board's most severe sanction, which cancels a doctor's license to practise. So let's first **filter** the data to look at those actions:

```{r}
# filter for license revocations only
revoked <- ca_discipline %>%
  filter(action_type == "Revoked")
  
datatable(revoked)
```

Some of these doctors are not actually based in California. Doctors can be licensed to practise in more than one state, and the Medical Board of California typically issues its own sanction if a doctor is disciplined in their home state. So let's now **filter** the data to look only at doctors with revoked licenses who were based in California, and **sort** them by city.

```{r}
# filter for license revocations by doctors based in California, and sort by city
revoked_ca <- ca_discipline %>%
  filter(action_type == "Revoked"
         & state == "CA") %>%
  arrange(city)

datatable(revoked_ca)
```

Here, the `filter` combines two conditions with `&`. That means that both have to be met for the data to be included. Then `arrange` **sorts** the data, which for a text variable will be in alphabetical order. If you wanted to sort in reverse alphabetical order, the code would be: `(arrange(desc(city))`.

#### Group and summarize data

Next we will **group** and **summarize** data by counting disciplinary actions by year and by month. But before doing that, we need to use the `year` and `month` functions from the **lubridate** package to extract the year and month from `action_date`.

```{r}
# extract year and month from action_date
ca_discipline <- ca_discipline %>%
  mutate(year = year(action_date),
         month = month(action_date))
```

Now we have two extra columns in the data, giving the year and the month as a number.

Previously we used the **dplyr** `mutate` function to modify an existing variable. Here we used it to create new variables. You can create or modify multiple variables in the same `mutate` function, separating each one by commas. Notice the use of `=` to make a variable equal to the output of code, within the `mutate` function.

Now we can calculate the number of license revocations for doctors based in California by year:

```{r}
# license revokations for doctors based in California, by year
revoked_ca_year <- ca_discipline %>%
  filter(action_type == "Revoked" 
         & state == "CA") %>%
  group_by(year) %>%
  summarize(revocations = n())
```

The code first **filters** the data, as before, then **groups** by the new variable `year` using `group_by`, then **summarizes** by counting the number of records for each year using `summarize`. The last function creates a new variable called `revocations` from `n()`, which is a count of the rows in the data for each year.

This should be the result:

```{r}
datatable(revoked_ca_year)
```

Looking at this result and the raw `ca_discipline` data, we only have partial data for 2008. So if we want to count the number of license revocations by month over all the years, we should first **filter** out the data for 2008, which will otherwise skew the result.

```{r}
# license revokations for doctors based in California, by month
revoked_ca_month <- ca_discipline %>%
  filter(action_type == "Revoked" 
         & state == "CA"
         & year >= 2009) %>%
  group_by(month) %>%
  summarize(revocations = n())

datatable(revoked_ca_month)
```


Notice how we used `>=` to filter for data where the year was 2009 or greater. The following code will achieve the same result:

```{r}
# license revocations for doctors based in California, by month
revoked_ca_month <- ca_discipline %>%
  filter(action_type == "Revoked" 
         & state == "CA"
         & year != 2008) %>%
  group_by(month) %>%
  summarize(revocations = n())
```

We can **group** and **summarize** by more than one variable at a time. The following code counts the number of actions of all types, by month and year. Again, we will first **filter** out the incomplete data for 2008.

```{r}
# disciplinary actions for doctors in California by year and month, from 2009 to 2017
actions_year_month <- ca_discipline %>%
  filter(state == "CA"
         & year >= 2009) %>%
  group_by(year, month) %>%
  summarize(actions = n()) %>%
  arrange(year, month)

datatable(actions_year_month)
```

Notice that the `group_by` and `arrange` functions **group** and **sort** the data, respectively, by two variables, separated by commas.

### Load and examine the Medicare opioid prescription data

Now we will work with the larger data on doctors in California prescribing opioids under Medicare Part D, learn how to **join** data in different data frames, and see how to make some simple charts to help make sense of your data.

Load and view the data:

```{r}
# load opioid prescription data
ca_opioids <- read_csv("ca_medicare_opioids.csv")

glimpse(ca_opioids)
```

#### Create a summary, showing the number of opioid prescriptions written by each doctor, the total cost of the opioids prescribed, and the cost per claim

```{r}
# Create a summary, showing the number of opioid prescriptions written by each provider, the total cost of the opioids prescribed, and the cost per claim
provider_summary <- ca_opioids %>% 
  group_by(npi,
           nppes_provider_last_org_name,
           nppes_provider_first_name,
           nppes_provider_city,
           specialty_description) %>%
  summarize(prescriptions = sum(total_claim_count),
            cost = sum(total_drug_cost)) %>%
  mutate(cost_per_prescription = cost/prescriptions) %>%
  arrange(desc(prescriptions))
```

Remember, when summarizing data, any `NA` values will cause the summary to be calculated as `NA`, unless you specify that nulls should be removed. If any of the cost or prescription values had been missing, for example, the `sum` functions would have to be written in this form: `sum(total_claim_count, na.rm = TRUE)`.

### Draw some charts from this summary data

While running data analyses, it's often useful to create charts to look at summaries of the data. Particularly useful are histograms, to examine a distribution, and scatterplots, to see how one variable varies with another.

To make simple charts, we will use **ggplot2**, which we explored in more depth in the previous quarter.

#### ggplot2 and the grammar of graphics

As a reminder, the "gg" in **ggplot2** stands for "[grammar of graphics](http://www.amazon.com/The-Grammar-Graphics-Statistics-Computing/dp/0387245448)," an approach to drawing charts devised by the statistician Leland Wilkinson. Rather than thinking in terms of finished charts like a scatter plot or a column chart, it starts by defining the coordinate system (usually the X and Y axes), maps data onto those coordinates, and then adds layers such as points, bars and so on. This is the logic behind ggplot2 code.

A reminder of **ggplot2** basics:

- `ggplot` This is the master function that creates a **ggplot2** chart.
- `aes` This function, named for "aesthetic mapping," is used whenever data values are mapped onto a chart. So it is used when you define which variables are plotted onto the X and Y axes, and also if you want to change the size or color of parts of the chart according to values for a variable.
- `geom` All of the functions that add layers to a chart start with `geom`, followed by an underscore, for example `geom_point` or `geom_bar`. The code in the parentheses for any `geom` layer styles the items in that layer, and can include `aes` mappings of values from data.
- `theme` This function modifies the appearance of elements of a plot, used, for example, to set size and font face for text, the position of a legend, and so on.
- `scale` Functions that begin with `scale`, followed by an underscore, are used to modify the way an `aes` mapping of data appears on a chart. They can change the axis range, for example, or specify a color palette to be used to encode values in the data.
- `+` is used each time you add a layer, a scale, a theme, or elements like axis labels and a title. After a `+` you can continue on the same line of code or move the next line. I usually write a new line after each `+`, which makes the code easier to follow.

First load **ggplot2** and the **[scales](https://cran.r-project.org/web/packages/scales/scales.pdf)** package, which allows numbers to be displayed as currency, with thousands separated by commas, and so on. Install **scales** if you don't have it already. We'll also load **plotly**, to allow us to explore the data on a chart using tooltips.

```{r}
library(ggplot2)
library(scales)
library(plotly)
```

#### Make a histogram of the prescriptions data

It's always a good idea to look at the distribution of any quantitative variable.

```{r}
# histogram of the costs data
ggplot(provider_summary, aes(x = prescriptions)) +
  geom_histogram()
```

This code first stated that the `doctor_summary` data would be used for the chart in the `ggplot` function, then specificied that values for the number of prescriptions should be mapped to the X, or horizontal axis. We didn't need to specify a variable for the Y axis for a histogram, because that is automatically a count of the number of records (here the number of doctors) in each bin.

The following code customizes the width of the bins, to 50 prescriptions, changes the X axis range between 0 and 3,000, and uses `labels` from the **scales** package to plot numbers with comma separators for thousands. As we learned last quarter, `theme_minimal` is one of the built-in themes in **ggplot2**, which gives a basic chart with grid lines on a white background.

```{r}
ggplot(provider_summary, aes(x = prescriptions)) +
  geom_histogram(binwidth = 50) +
  theme_minimal() +
  scale_x_continuous(limits = c(0,3000),
                     labels = comma) +
  scale_y_continuous(labels = comma)
```

This is a very skewed distribution. Viewing this, we would conclude that the mean would not be a good summary statistic for this data; if we want to compare any doctor's prescriptions of opioids to that for a "typical" doctor, it would be better to compare to the median number of prescriptions. So let's calculate that number:

```{r}
median(provider_summary$prescriptions)
```

#### Make a scatterplot of prescriptions and costs data

Scatterplots are good for seeing how one variable relates to one another, and also for spotting outliers in data. Here we would expect the total cost of the drugs prescribed to increase with the total number of prescriptions. A scatterplot allows us to see whether all the providers are closely clustered around the trend line, or whether there are some interesting outliers.

This is a good case for making a chart interactive, so we can hover over each point to find out who any provider is.

```{r}
#### Make a scatterplot of prescriptions and costs data
scatterplot <- ggplot(provider_summary, aes(x = prescriptions, 
                                            y = cost,
                                            text = paste0("<b>Name: </b>",nppes_provider_first_name," ",nppes_provider_last_org_name,"<br>",
                                                          "<b>Location: </b>",nppes_provider_city,"<br>",
                                                          "<b>Specialty: </b>",specialty_description,"<br>",
                                                          "<b>NPI: </b>",npi))) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = lm) +
  theme_minimal() +
  scale_x_continuous(labels = comma) +
  scale_y_continuous(labels = dollar)

ggplotly(scatterplot, tooltip = "text") %>%
  config(displayModeBar = FALSE)
```

In this code, we mapped the number of `prescriptions` to the X axis and their total `cost` to the Y or vertical axis. We then used `geom_point` to add the points, and made them partially transparent using `alpha = 0.3` (0 would be completely transparent; 1 would be completely opaque). `geom_smooth` added a trend line through the data; `method = lm` ensured that this is fitted with a "linear model," giving a straight line. `labels = dollar` formatted the Y axis so the numbers were shown in dollars.

Remember the words of caution: There may be some story leads here, but we would have to find out more about each doctor's practise and patients to determine whether outliers in terms of prescriptions and costs are justified.

### Join data to find doctors who prescribed opioids who have also been subject to disciplinary actions.

We will join the `provider_summary` data to the disciplinary actions data to see which of the doctors prescribing opioids have been in trouble with the Medical Board of California. The unique identifier in the prescriber data is their National Provider Identifier, whereas the unique identifier in the disciplinary data are the doctors' California medical license numbers. So we also need to load the file `npi_license.csv`, which relates these two identifiers to one another.

```{r}
# load data
npi_license <- read_csv("npi_license.csv")
```

First we will join these two files to one another:

```{r}
# join those two data frames
ca_discipline_npi <- left_join(ca_discipline, npi_license)
```

By default, **dplyr** will join by any variables with matching names, but you can also specify the variables on which to join. So this will achieve the same result:

```{r}
# join those two data frames
ca_discipline_npi <- left_join(ca_discipline, npi_license, by = "license")

datatable(ca_discipline_npi)
```

The first few rows reveal some of the limitations of the `npi_license` data. Some of the doctors have not been joined to an `npi` code, while others appear to have more than one `npi` code (this is why there are 9,680 rows in the joined data, and only 7,651 in the `ca_discipline` data).

Having made this join, we can now join to the `provider_summary` data:

```{r}
# join disciplinary action data to the opioid prescription data
provider_summary_actions <- inner_join(provider_summary, ca_discipline_npi, by = "npi") %>%
  arrange(desc(prescriptions))

datatable(provider_summary_actions)
```

Individual doctors may appear multiple times in the data if they have more than one disciplinary action on their file, because a join will be made to each individual action.

We can search for the disciplinary records and look at public documents at [this site](https://www.breeze.ca.gov/).

The second doctor on our list, Naga Thota, surrendered his California medical license on March 22, 2017. [This public document](http://www2.mbc.ca.gov/BreezePDL/document.aspx?path=%5cDIDOCS%5c20170315%5cDMRAAAFC8%5c&did=AAAFC170315180454982.DID) reveals that he was charged with prescribing excessive quantities of opioids to an addicted patient with whom he also engaged in a sexual relationship. In addition to disciplinary action from the Medical Board of California, he faced a criminal investigation by the federal Drug Enforcement Adminstration. In March 2017 [sentenced to two years](http://www.sandiegouniontribune.com/news/courts/sd-me-thota-sentencing-20170306-story.html), he was sentenced to two years in a federal prison.

The prescriptions data reveal that, prior to his conviction and the loss of his medical license, Thota was also one of the largest prescribers of opioids in the state of California.

There are likely to be other story leads from this join, which you could follow up through public records and other reporting.

We've already seen that the `npi_license` data has some limitations. Another (less reliable) way to join the data would be by using the doctors' first names, last names, and cities. First, however, we need to convert those variables in the `ca_discipline_npi` data to upper case, so they match the format of the `ca_opioids` data.

```{r}
# change case of variables to be used in the join
ca_discipline_npi <- ca_discipline_npi %>%
  mutate(last_name = toupper(last_name),
         first_name = toupper(first_name),
         city = toupper(city))

# join disciplinary action data to the opioid prescription data
provider_summary_actions_2 <- inner_join(provider_summary, ca_discipline_npi, by = c("nppes_provider_last_org_name" = "last_name", 
                                                                                 "nppes_provider_first_name" = "first_name",
                                                                                 "nppes_provider_city" = "city")) %>%
  arrange(desc(prescriptions))
```

Notice how the code is written to define the variables to be used for a custom join.

Now we can use an `anti_join` to see if this join picked up any doctors who were missing from the first.

```{r}
# join disciplinary action data to the opioid prescription data
provider_summary_actions_extra <- anti_join(provider_summary_actions_2, provider_summary_actions)

datatable(provider_summary_actions_extra)
```


Whenever making an join in this way, it crucial to verify through further reporting that the individuals matched are the same person. For example, there are several doctors called Peter Lee in Los Angeles. The one with the disciplinary record is [Peter Geon Lee](https://search.dca.ca.gov/details/8002/G/84673/aebacf63f2746d6ce6ac729ebbbe54f5), a plastic surgeon. This is not [Peter Chong Lee](https://search.dca.ca.gov/details/8002/A/47846/e516806f7ff2db771d5b14b7d175c755), the physical medicine and rehabilitation specialist, with the NPI of `1063508786`, listed in the `ca_opioids` data. There may well be other false matches in the data, joined in this way.

### Further reading

**[Introduction to dplyr](https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html)**

**[RStudio Data Wrangling Cheet Sheet](https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf)**
Also introduces the [**tidyr**](https://blog.rstudio.org/2014/07/22/introducing-tidyr/) package, which can manage wide-to-long transformations, and text-to-columns splits, among other data manipulations.

**[Stack Overflow](http://stackoverflow.com/)**
For any work involving code, this question-and-answer site is a great resource for when you get stuck, to see how others have solved similar problems. Search the site, or [browse R questions](http://stackoverflow.com/questions/tagged/r)

[***R Graphics Cookbook***](http://www.amazon.com/R-Graphics-Cookbook-Winston-Chang/dp/1449316956) by Winston Chang
(Chang also has [a helpful website](http://www.cookbook-r.com/) with much of the same information, available for free.)

[**RStudio ggplot2 Cheat Sheet**](https://www.rstudio.com/wp-content/uploads/2015/08/ggplot2-cheatsheet.pdf)

[**ggplot2 documentation**](http://ggplot2.tidyverse.org/reference/)





